---
title: "Landscape_Indices"
author: "Maite"
format: html
editor: visual
theme: flatly
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=F, warning = F)
```

```{r}
library(here)
library(sf)
library(terra)
library(raster)
library(mapview)
```

## Landscape Indices

Trying some indices with the DLM data.

Ideas:

-   How much grassland surrounds the patch? Area

-   How is the vicinity of the grassland?

    -   E.g. forest or city?

    -   In which proportions and distances? 

    -   perhaps also directions and elevations of the neighboring grassland patches (for aerial and fluvial transport)

-   

Using Thüringen as an example:

```{r}
#| results: 'hide'
#| fig.keep: 'all'
dp <- "../data/"
sap <- here(dp, "/studyarea")

geo_exp <- st_read(here(sap, "all3explos.gpkg"))
plots_exp <- st_read(here(sap, "experiementalplots.gpkg"))
lsc_all <- st_read(here(sap, "landscape_all.gpkg"))


```

```{r eval=F}
#| warning: false
#| eval: false
dlm_th <- st_read(here("../data/Basisdaten/DLM/Thueringen/veg01_f.shp"))
dlm_th <- st_transform(dlm_th, crs(lsc_all))
dlm_th_lsc <- st_intersection(lsc_all[2,], dlm_th)
st_write(dlm_th_lsc, here("../data/Basisdaten/DLM/Thueringen/veg01_f_proj_cut.shp"))

```

```{r}
#| results: 'hide'
#| fig.keep: 'all'
dlm_th <- st_read(here("../data/Basisdaten/DLM/Thueringen/veg01_f_proj_cut.shp"))

```

For now only using the Exploratory Area of the DLM.

```{r}
dlm_th <- st_intersection(geo_exp[2,], dlm_th)

dlm_th_grass <- dlm_th[dlm_th$VEG=="1020",]

mapview(dlm_th_grass, map.types = "Esri.WorldImagery", col.regions="#139404")+
  mapview(lsc_all[2,], alpha=0)+ 
  mapview(plots_exp[plots_exp$explrtr == "HAI" & plots_exp$type == "grassland",])


```

Which grassland geometries contain plots: 
```{r}
has_plot <- st_contains(dlm_th_grass, plots_exp)
has_plot <- unlist(as.character(has_plot))
has_plot[has_plot == "integer(0)"] <- NA
dlm_th_grass$has_plot <- !is.na(has_plot)

mapview(dlm_th_grass, zcol="has_plot")
```

Probably best to unify patches:


Unify adjacent grassland patches: 
Helpful: https://github.com/rstudio/cheatsheets/blob/main/sf.pdf
https://github.com/r-spatial/sf/issues/290 

run st_touches and loop st_union until all touching are dissolved? 

```{r}
library(tidyr)
library(dplyr)


# start loop here and end when touch is completely empty 

touch <- st_touches(dlm_th_grass)


touchidlist <- lapply(seq(touch), function(i){
  c(i, which(df[,i] != 0))
})

dis <- lapply(seq(touchidlist), function(i){
  summarise(dlm_th_grass[touchidlist[[i]], ], 
          m=mean(AREA))
})

dis_col <- do.call(rbind, dis)
dis_col

# loop ends here 

# just checking
mapview(dis_col, alpha.regions=0, lwd=2, color="red")+
  mapview(dlm_th_grass, alpha.regions=0, lwd=1)

```

That's not it.. that way we loose all information 
```{r}
dlm_th_grass_union <- st_union(dlm_th_grass, by_feature=TRUE)

mapview(dlm_th_grass, col.regions="#139404", map.types = "Esri.WorldImagery")+
  mapview(dlm_th_grass_union)+
  mapview(plots_exp, query.position = mapviewGetOption("query.position"))
```


### Calculate area grassland around patch
```{r}
gli <- st_intersection( dlm_th_grass_union, plots_exp)
```

### Calculate proportion of other DLM types, their distance and orientation respective to the plot
```{r}

```



Trying landscapemetrics package: 
```{r}
library(landscapemetrics)
```


## Sources:

https://gistbok.ucgis.org/bok-topics/landscape-metrics#:\~:text=Landscape%20metrics%E2%80%94also%20known%20as,within%20a%20defined%20geographic%20area.
