---
title: "comparing soil property data"
format: html
editor: visual
theme: flatly
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=F, warning = F)
```

## Aim

This document compares gridded soil data to the data gathered at the EPs.

For this, I'm using the dataset ID 31210 from the Soil abiotic (core) project's Soil Sampling Campaign 21 from BExis [Soil sampling campaign 2021, all experimental plots (EP), 0-10 cm - carbon and nitrogen concentrations](https://www.bexis.uni-jena.de/ddm/data/Showdata/31210) with the following variables:

```{r}


library(here)
library(sf)
library(mapview)
library(dplyr)
library(terra)
library(ggplot2)
```

```{r}
#| results: 'hide'
#| fig.keep: 'all'

sdat <- read.csv(here("../data/BExis/31210_6_Dataset/31210_6_data.csv"))
d_geo <- st_read(here("../data/16826_4.gpkg"))


correct_ep_PlotID <- function(data, epID_Colname="EP_Plotid"){ # add 0 where ID only contains 4 characters in d
  id <- data[epID_Colname][,1]
  id[nchar(id) == 4] <- paste0(substring(id[nchar(id) == 4], 1,3), "0", 
                                             substring(id[nchar(id) == 4], 4))
  data[epID_Colname] <- id
  return(data)
}

sdat <- correct_ep_PlotID(sdat)

# select only grassland 
sdat <- sdat |>
  filter(Type == "G")

d_geo <- d_geo |>
  select(c("shannon", "biomass_g", "ep"))

sdat_geo <- merge(d_geo, sdat, by.x="ep", by.y="EP_Plotid")
```

```{r}
names(sdat)[5:length(names(sdat))]
mapview(sdat_geo, zcol="CN_ratio")

```

There is also a pH dataset from the abiotic soil core team with the dataset ID 31074 [Soil sampling campaign 2021, all experimental plots (EP), 0-10 cm - soil pH](https://www.bexis.uni-jena.de/ddm/data/Showdata/31074). "The soil pH was measured in a weak (0.01 M) calcium chloride solution using a pH meter."

```{r}

pHdat <- read.csv(here("../data/BExis/31074_8_Dataset/31074_8_data.csv"))
pHdat <- correct_ep_PlotID(pHdat)

# select only grassland 
pHdat <- pHdat |>
  filter(Type == "G") |>
  select(c( "EP_Plotid","pH_1", "pH_2"))

ph_sdat_geo <- merge(sdat_geo, pHdat, by.x="ep", by.y="EP_Plotid")

```

```{r}
corrdat <- ph_sdat_geo |>
  as.data.frame() |>
  select(c(2,3,7:13))

cmt <- psych::corr.test(corrdat, method="pearson")
corrplot::corrplot(cmt$r, p.mat= cmt$p,
                   method="number", 
                   type="lower", 
                   diag = F, 
                   tl.col="black")



```

Moreover, there is the [MinSoil 2017 Soil Sampling](https://www.bexis.uni-jena.de/ddm/data/Showdata/23886) from the Soil abiotic (core) team containing sand, silt and clay.

```{r}
texdat <- read.csv(here("../data/BExis/14686_10_Dataset/14686_10_data.csv"))
texdat <- correct_ep_PlotID(texdat)

texdat <- texdat[texdat$Type=="G",]

```

## ESDAC

Get the ESDAC chemical data:

-   pH (measured in H2O)

-   pH (n CaCl2 0.01 M solution)

-   Cation Exchange Capacity (CEC)

-   Calcium carbonates (CaCO3)

-   C:N ratio Nitrogen (N)

-   Phosphorus (P)

-   Potassium (K)

```{r}
#| fig-height: 9
path <- here("../data/Basisdaten/soil/ESDAC/chemical/")
soil_chem <- rast(paste0(path, "/soil_chem_ger.grd"))
names(soil_chem)

EP_chem <- extract(soil_chem, ph_sdat_geo)

sdat_chem <- cbind(ph_sdat_geo, EP_chem)

sdat_chem_corrdat <- sdat_chem |>
  as.data.frame() |>
  select(c(2,3,7:13, 15:23))

cmt <- psych::corr.test(sdat_chem_corrdat, method="pearson")
corrplot::corrplot(cmt$r, p.mat= cmt$p,
                   method="number", 
                   type="lower", 
                   diag = F, 
                   tl.col="black")

```

y axis contains data from ESDAC dataset (500m resolution), x axis contains data gathered by BE core team:

```{r}
#| layout-ncol:  2

r <- cor(sdat_chem$CN, sdat_chem$CN_ratio, use = "pairwise.complete")
t <- "CN ratio"
ggplot(sdat_chem, aes(CN_ratio, CN))+
  geom_point()+
  ggtitle(t, subtitle  = paste0("r = ", round(r, digits=2)))+
  geom_abline(slope=1, lty="dotted")+
  xlab(paste0(t, " BEx"))+
  ylab(paste0(t, " ESDAC"))+
  theme_minimal()

r <- cor(sdat_chem$Total_N, sdat_chem$N, use = "pairwise.complete")
t <- "N"
ggplot(sdat_chem, aes(Total_N, N))+
  geom_point()+
  ggtitle(t, subtitle  = paste0("r = ", round(r, digits=2)))+
  geom_abline(slope=1, lty="dotted")+
  theme_minimal()


r <- cor(sdat_chem$pH_1, sdat_chem$pH_H2O_CaCl, use = "pairwise.complete")
t <- "1st pH_H20_CaCl"
ggplot(sdat_chem, aes(pH_1, pH_H2O_CaCl))+
  geom_point()+
  ggtitle(t, subtitle  = paste0("r = ", round(r, digits=2)))+
  geom_abline(slope=1, lty="dotted")+
  theme_minimal()

r <- cor(sdat_chem$pH_2, sdat_chem$pH_H2O_CaCl, use = "pairwise.complete")
t <- "2nd measurement pH_H20_CaCl"
ggplot(sdat_chem, aes(pH_2, pH_H2O_CaCl))+
  geom_point()+
  ggtitle(t, subtitle  = paste0("r = ", round(r, digits=2)))+
  geom_abline(slope=1, lty="dotted")+
  theme_minimal()



```

## SoilGrid.org data

Downloaded N, pH and SOC, each with uncertainty for the three explo areas.

```{r}
folder <- here("../data/Basisdaten/soil/SoilGrids/same_size")

heg <- rast(paste0(folder, "/HEG.grd"))
seg <- rast(paste0(folder, "/SEG.grd"))
alb <- rast(paste0(folder, "/ALB.grd"))


plot(seg$SEG_SOC)
plot(heg$HEG_SOC)
plot(alb$ALB_SOC)

extraction_template <- st_transform(ph_sdat_geo, crs(heg))

extract_SoilGrids <- function(x){
    out <- extract(x, extraction_template)
    out <- cbind(extraction_template$ep, out)
    out <- out[!is.na(out[,3]),]
    names(out) <- paste0("SG_", c( "ep", "ID", "N", "N_unc", "pH", "pH_unc", "SOC", "SOC_unc"))
    return(out)
}

heg_out <- extract_SoilGrids(heg)
seg_out <- extract_SoilGrids(seg)
alb_out <- extract_SoilGrids(alb)

extraction_results <- rbind(heg_out, seg_out, alb_out)

# applying correction factors
extraction_results$SG_SOC <- extraction_results$SG_SOC / 10
extraction_results$SG_N <- extraction_results$SG_N / 100
extraction_results$SG_pH <- extraction_results$SG_pH / 10


sdat_SG <- merge(ph_sdat_geo, extraction_results, by.x="ep", by.y="SG_ep")

```

```{r}
#| layout-ncol:  2

r <- cor(sdat_SG$SG_N, sdat_SG$Total_N, use = "pairwise.complete")
t <- "N"
ggplot(sdat_SG, aes(Total_N, SG_N, col=SG_N_unc))+
  geom_point()+
  scale_colour_viridis_c()+
  ggtitle(t, subtitle  = paste0("r = ", round(r, digits=2)))+
  geom_abline(slope=1, lty="dotted")+
  xlab(paste0(t, " BEx"))+
  ylab(paste0(t, " SoilGrids"))+
  theme_minimal()


r <- cor(sdat_SG$Organic_C, sdat_SG$SG_SOC, use = "pairwise.complete")
t <- "SOC"
ggplot(sdat_SG, aes(Organic_C, SG_SOC, col=SG_SOC_unc))+
  geom_point()+
  scale_colour_viridis_c()+
  ggtitle(t, subtitle  = paste0("r = ", round(r, digits=2)))+
  geom_abline(slope=1, lty="dotted")+
  xlab(paste0(t, " BEx"))+
  ylab(paste0(t, " SoilGrids"))+
  theme_minimal()

r <- cor(sdat_SG$SG_pH, sdat_SG$pH_1, use = "pairwise.complete")
t <- "pH 1"
ggplot(sdat_SG, aes(pH_1, SG_pH, col=SG_pH_unc))+
  geom_point()+
  scale_colour_viridis_c()+
  ggtitle(t, subtitle  = paste0("r = ", round(r, digits=2)))+
  geom_abline(slope=1, lty="dotted")+
  xlab(paste0(t, " BEx"))+
  ylab(paste0(t, " SoilGrids"))+
  theme_minimal()

r <- cor(sdat_SG$SG_pH, sdat_SG$pH_2, use = "pairwise.complete")
t <- "pH 2"
ggplot(sdat_SG, aes(pH_2, SG_pH, col=SG_pH_unc))+
  geom_point()+
  scale_colour_viridis_c()+
  ggtitle(t, subtitle  = paste0("r = ", round(r, digits=2)))+
  geom_abline(slope=1, lty="dotted")+
  xlab(paste0(t, " BEx"))+
  ylab(paste0(t, " SoilGrids"))+
  theme_minimal()

```

pH doesn't look too good.. At the same time the correlation between the two BE pH measurements (below) is only about 0.5, so perhaps we might provey the pH signal better by pH predictors like soil depth, upslope area and bedrock or something like that? This could be a good use case for a latent variable in SEM (?)

```{r}
cor.test(sdat_SG$pH_2, sdat_SG$pH_1, use = "pairwise.complete")
```

In terms of N and SOC, SoilGrids totally fails for the Schorfheide Plots.

```{r}
r <- cor(sdat_SG$SG_N, sdat_SG$Total_N, use = "pairwise.complete")
t <- "N"
ggplot(sdat_SG, aes(Total_N, SG_N, col=SG_N_unc, label=ep))+
  geom_point()+geom_text(hjust=0, vjust=0)+
  scale_colour_viridis_c()+
  ggtitle(t, subtitle  = paste0("r = ", round(r, digits=2)))+
  geom_abline(slope=1, lty="dotted")+
  xlab(paste0(t, " BEx"))+
  ylab(paste0(t, " SoilGrids"))+
  theme_minimal()


r <- cor(sdat_SG$Organic_C, sdat_SG$SG_SOC, use = "pairwise.complete")
t <- "SOC"
ggplot(sdat_SG, aes(Organic_C, SG_SOC, col=SG_SOC_unc, label=ep))+
  geom_point()+geom_text(hjust=0, vjust=0)+
  scale_colour_viridis_c()+
  ggtitle(t, subtitle  = paste0("r = ", round(r, digits=2)))+
  geom_abline(slope=1, lty="dotted")+
  xlab(paste0(t, " BEx"))+
  ylab(paste0(t, " SoilGrids"))+
  theme_minimal()

r <- cor(sdat_SG$SG_pH, sdat_SG$pH_1, use = "pairwise.complete")
t <- "pH 1"
ggplot(sdat_SG, aes(pH_1, SG_pH, col=SG_pH_unc, label=ep))+
  geom_point()+geom_text(hjust=0, vjust=0)+
  scale_colour_viridis_c()+
  ggtitle(t, subtitle  = paste0("r = ", round(r, digits=2)))+
  geom_abline(slope=1, lty="dotted")+
  xlab(paste0(t, " BEx"))+
  ylab(paste0(t, " SoilGrids"))+
  theme_minimal()

r <- cor(sdat_SG$SG_pH, sdat_SG$pH_2, use = "pairwise.complete")
t <- "pH 2"
ggplot(sdat_SG, aes(pH_2, SG_pH, col=SG_pH_unc, label=ep))+
  geom_point()+geom_text(hjust=0, vjust=0)+
  scale_colour_viridis_c()+
  ggtitle(t, subtitle  = paste0("r = ", round(r, digits=2)))+
  geom_abline(slope=1, lty="dotted")+
  xlab(paste0(t, " BEx"))+
  ylab(paste0(t, " SoilGrids"))+
  theme_minimal()

```

```{r}
path <- here("../data/Basisdaten/soil/ESDAC/physical/")
soil_phys <- rast(paste0(path, "/soil_phys_ger.grd"))
phys_stars <- stars::st_as_stars(soil_phys)


stars::st_as_stars()

mapview(phys_stars[,,,7])+mapview(ph_sdat_geo)

plot(soil_phys)


lui <- read.csv(list.files(here("../data/LUI/LuiData/"), pattern=".txt$", full.names=T))

# fix some things in LUI dataset to merge correctly 
lui$year <- substring(lui$YEAR, 12,15)

idshort <- lui$PLOTID[nchar(lui$PLOTID) == 4] 
idshortfix <- paste0(substring(idshort, 1,3), "0", substring(idshort, 4,4))
lui$PLOTID[nchar(lui$PLOTID) == 4]  <- idshortfix

d_lui <- merge(d_geo, lui, by.x="ep", by.y="PLOTID")

ml <- d_lui |> as.data.frame() |> 
  select(-"geometry") |>
  group_by(ep) |>
  reframe(mean_LUI = mean(LUI))

mm <- d_lui |> as.data.frame() |> 
  select(-"geometry") |>
  group_by(ep) |>
  reframe(mean_mowing = mean(M_STD))

sdat_lui <- merge(sdat_geo, ml, by="ep")
sdat_lui <- merge(sdat_lui, mm, by="ep")

r <- cor(sdat_lui$mean_LUI, sdat_lui$Organic_C, use = "pairwise.complete")
ggplot(sdat_lui, aes(mean_LUI, Organic_C, label=ep))+
  geom_point()+geom_text(hjust=0, vjust=0)+
  scale_colour_viridis_c()+
  ggtitle("LUI and Organic C", subtitle  = paste0("r = ", round(r, digits=2)))+
  theme_minimal()


r <- cor(sdat_lui$mean_LUI, sdat_lui$Organic_C, use = "pairwise.complete")
ggplot(sdat_lui, aes(mean_mowing, Organic_C, label=ep))+
  geom_point()+geom_text(hjust=0, vjust=0)+
  scale_colour_viridis_c()+
  ggtitle("Mowing and Organic C", subtitle  = paste0("r = ", round(r, digits=2)))+
  theme_minimal()
```

Troubleshooting:

Why is SEG falling out of the expected relation that much?

Bodenart, Bodenfeuchte (weniger umgesetzt, mehr gespeichert), LUI could also be (grazing, mowing)

Sand content - should have a negative relation, though...
