---
title: "SoilGrids"
author: "Maite"
format: html
editor: visual
theme: flatly
toc: true
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=F, warning = F)
```

## Get data for germany

R download VRT (Virtual Raster) as specified on [this website](https://git.wur.nl/isric/soilgrids/soilgrids.notebooks/-/blob/master/markdown/webdav_from_R.md). Data structure online is [here](https://files.isric.org/soilgrids/latest/data/) and the explanation of the short names can be found under "Properties" [here](https://www.isric.org/explore/soilgrids/faq-soilgrids).

Interesting for now:

````         
```         
| name     | description                                      | mapped units | conversion factor to conventional units | conventional units |
|-------------|-------------|-------------|---------------------|-------------|
| clay     | Propotion clay particles in fine earth fraction  | g/kg         | 10                                      | g/100g (%)         |
| nitrogen | total N                                          | cg/kg        | 100                                     | g/kg               |
| phh2o    | Soil pH                                          | pHx10        | 10                                      | pH                 |
| sand     | proportion sand particles in fine earth fraction | g/kg         | 10                                      | g/100g (%)         |
| silt     | prop silt particles                              | g/kg         | 10                                      | g/100g (%)         |
| soc      | soil org c in fine earth frac                    | dg/kg        | 10                                      | g/kg               |
| ocd      | Organic carbon density                           | hg/m\^3      | 10                                      | kg/m\^3            |
| ocs      | organic carbon stocks                            | t/ha         | 10                                      | kg/m\^2            |
```
````

```{r}
#| results: 'hide'
#| fig.keep: 'all'

library(terra)
library(gdalUtilities)
library(sf)
library(here)

dp <- "../data/"
sap <- here(dp, "/studyarea")
cp <- here("../data/Basisdaten/landscape/Copernicus/")
germany <- st_read(paste0(sap, "germany.gpkg"))

bb <- raster::extent(germany)[c(1,3,2,4)] #xmin,ymin,xmax,ymax
crs <- crs(germany)


sg_url="/vsicurl?max_retry=3&retry_delay=1&list_dir=no&url=https://files.isric.org/soilgrids/latest/data/"

outDir <- here("../data/Basisdaten/soil/SoilGrids/")

```

```{r}
#| eval: false

# check whether bounding box needs to be put together differently 

# get all possible combinations of 1:4
gr <- as.data.frame( expand.grid(1:4, 1:4, 1:4, 1:4))
all_in <- apply(gr, 1, function(x) all(c(1,2,3,4) %in% x) )
gr_all_in <- gr[all_in,]


check_all <- lapply(seq(nrow(gr_all_in)), function(i){
  
    v <- unlist(gr_all_in[i,])
  
    bb <- raster::extent(germany)[v]
    
    tryCatch(gdal_translate(paste0(sg_url,'ocs/ocs_0-30cm_mean.vrt'),
        paste0(outDir, "/GER_ocs_mean.tif"),
        tr=c(250,250),
        projwin=bb,
        projwin_srs=crs), 
        error=function(e) {"doesn't work"})

})

any(check_all != "doesn't work")


```

Didn't work. Downloaded manually on SoilGrids.org:

```{r}
#| eval: false

#### PREPARE RASTERS ##################################

# downloaded pH first, that's smaller extent 
folder <- here("../data/Basisdaten/soil/SoilGrids/same_size")


# HEG
hegfiles <- list.files(here("../data/Basisdaten/soil/SoilGrids/"), pattern="HEG", full.names = T)

hegfileminex <- rast(list.files(here("../data/Basisdaten/soil/SoilGrids/"), pattern="HEG_pH.tif", full.names = T))

heg <- lapply(seq(hegfiles), function(i){
  r <- rast(hegfiles[[i]])
  r <- crop(r, hegfileminex)
  ext(r)
  names(r)
  writeRaster(r, paste0(here("../data/Basisdaten/soil/SoilGrids/same_size"), "/", names(r), ".grd"), overwrite=T)
})

heg_matchedfiles <- list.files(folder, pattern=".grd$", full.names=T)
heg_matchedfiles <- heg_matchedfiles[grepl("HEG", heg_matchedfiles)]
heg_matchedfiles <- heg_matchedfiles[!grepl("SoilGroup", heg_matchedfiles)]

heg <- rast(heg_matchedfiles)
writeRaster(heg, paste0(here("../data/Basisdaten/soil/SoilGrids/same_size"), "/HEG.grd"), overwrite=T)


# for SEG 
segfiles <- list.files(here("../data/Basisdaten/soil/SoilGrids/"), pattern="SEG", full.names = T)
segfiles <- segfiles[!grepl("SoilGroup", segfiles)]

SEG <- rast(segfiles)

writeRaster(SEG, paste0(here("../data/Basisdaten/soil/SoilGrids/same_size"), "/SEG.grd"), overwrite=T)

# ALB
albfiles <- list.files(here("../data/Basisdaten/soil/SoilGrids/"), pattern="ALB", full.names = T)
albfiles <- albfiles[!grepl("SoilGroup", albfiles)]

ALB <- rast(albfiles)
writeRaster(ALB, paste0(here("../data/Basisdaten/soil/SoilGrids/same_size"), "/ALB.grd"), overwrite=T)


```

Downloaded N, pH and SOC, each with uncertainty for the three explo areas.

```{r}
folder <- here("../data/Basisdaten/soil/SoilGrids/same_size")

heg <- rast(paste0(folder, "/HEG.grd"))
seg <- rast(paste0(folder, "/SEG.grd"))
alb <- rast(paste0(folder, "/ALB.grd"))


names(alb)
```
